Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering

### 2016-07-11 paper/structureScene.tex
ref in intro

The proposed approach is motivated in Section \ref{sec:motivations} and a brief review the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}. Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively. Experiments carried out in an unsupervised setting, \ie acoustic scene similarity retrieval (ASSR) and in a supervised setting, \ie acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:ExpEval}.

\ml{blurb about results}

@@

The same dichotomy applies to the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
In the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag}, which consists in modeling audio by high-level summary statistics computed from a set of local features.
As it enjoys a high genericity, this approach is found in many audio classification tasks besides ASC.
Yet, the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be perform equivalently compared to a mere averaging of the features  \cite{lagrange:hal-01082501}.
Indeed, the typical morphology of acoustic scenes is of a "skeleton of events on a bed of textures" \cite{nelken_ear_2013}, \ie few impulsive sounds superimposed upon a stationary background.
Therefore, it is arguably more beneficial to focus on the temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.
Interestingly, this statement is validated by auditory psychology as well as sound synthesis based  on summary statistics \cite{mcdermott2013summary}.

@@

\vl{missing transition}

@@

The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function (RBF) kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
\begin{equation}
K^c_{ij} = exp\left( - \dfrac{\Vert c_i - c_j \Vert^2}{\Vert c_i - c_N \Vert \Vert c_i - c_N \Vert}
\end{equation}

where $c_N$ is the $N^{\mathrm{th}}$ nearest neihgbor of the centroid $c_i$, and $\Vert \cdot \Vert$ denotes the Euclidean norm.

@@

\section{Experiments}
\label{sec:ExpEval}

blurb

\subsection{Features}

@@

\section{Results}

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{gfx/unsupervised_test3.eps}
\caption{ASS: $p@k$ obtained for MFCCs and scattering with logarithmic compression}
\label{fig:ASS_1}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{gfx/unsupervised_test2.eps}
\caption{ASS: $p@k$ obtained for scattering with and without logarithmic compression (w/o log)}
\label{fig:ASS_2}
\end{center}
\end{figure}

@@

This section present evaluation results for the ASS settings. The $p@k$ for different settings are showed in the Figures~\ref{fig:ASS_1} and \ref{fig:ASS_2}.

@@

\begin{table}
\begin{center}
\begin{tabular}{llll}
                    & log-scattering & scattering & MFCCs \\
             \hline
ei (linear)  & 66 (11)        & 62 (6)  & 46 (8)    \\
ei (RBF)     & 60 (7)         & 57 (8)  & 60 (8)    \\
li (linear)  & 75 (5)         & 70 (8)  & 63 (8)   \\
li (RBF)     & 78 (6)         & 73 (8)  & 72 (10)   \\
ob-c         & 66 (7)         & 58 (6)  & 56 (9)   \\
ob-a         & 61 (7)         & 54 (8)  & 61 (9)   \\
ob-w         & 54 (11)        & 43 (10) & 51 (15)   \\
\end{tabular}
\caption{FYW: best results for object-based approach with majority vote (on the centroids / on the centroids + on the frames) is circa 75\%}
\end{center}
\end{table}

@@

\vl{Results: DCASE 2016}

>>>>

The proposed approach is motivated in Section \ref{sec:motivations} and a brief review the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}. Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively. Experiments carried out in an unsupervised setting, \ie acoustic scene similarity retrieval (ASSR) and in a supervised setting, \ie acoustic scene classification (ASC) are described in \ref{sec:experiments}. Results are subsequently reported in Section \ref{sec:results}.

@@

The same dichotomy applies to the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
In the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag}, which consists in modeling audio by high-level summary statistics computed from a set of local features.
As it enjoys a high genericity, this approach is found in many audio classification tasks besides ASC.
Yet, the specific implementation with gaussian mixture models (GMM) of Mel-frequency cepstral coefficients (MFCC) it is demonstrated to be perform equivalently compared to a mere averaging of the features  \cite{lagrange:hal-01082501}.
Indeed, the typical morphology of acoustic scenes is of a "skeleton of events on a bed of textures" \cite{nelken_ear_2013}, \ie few impulsive sounds superimposed upon a stationary background.
Therefore, it is arguably more beneficial to focus on the temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.
Interestingly, this statement is validated by auditory psychology as well as sound synthesis based  on summary statistics \cite{mcdermott2013summary}.

@@

In this paper, we shall demonstrate that the use of features based on the wavelet scattering together with object-based representation of the acoustic scene allows us to achieve interesting performance both in the ASSR and ASC paradigms.

@@

The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function (RBF) kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
\begin{equation}
\label{eq:kc}
\begin{split}
K^c_{ij} & =exp\left( - \dfrac{\parallel c_i - c_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
\sigma_{i} & =\parallel c_i - c_N \parallel
\end{split}
\end{equation}

with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$, and $\parallel \bullet \parallel$ the euclidean norm.

@@

\section{Experiments}
\label{sec:experiments}

\subsection{Feature design}

@@

\section{Results \label{sec:results}}

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{gfx/unsupervised_test3.eps}
\caption{ASSR: $p@k$ obtained for MFCCs and scattering with logarithmic compression}
\label{fig:ASS_1}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{gfx/unsupervised_test2.eps}
\caption{ASSR: $p@k$ obtained for scattering with and without logarithmic compression (w/o log)}
\label{fig:ASS_2}
\end{center}
\end{figure}

@@

This section present evaluation results for the ASSR settings. The $p@k$ for different settings are showed in the Figures~\ref{fig:ASS_1} and \ref{fig:ASS_2}.

@@

\begin{table}
\begin{center}
\begin{tabular}{llll}
                    & scattering & scattering & MFCCs \\
                    & w/ log comp.    & w/o log comp.   &      \\
             \hline
ei (linear)  & 66 (11)        & 62 (6)  & 46 (8)    \\
ei (RBF)     & 60 (7)         & 57 (8)  & 60 (8)    \\
li (linear)  & 75 (5)         & 70 (8)  & 63 (8)   \\
li (RBF)     & 78 (6)         & 73 (8)  & 72 (10)   \\
ob-c         & 66 (7)         & 58 (6)  & 56 (9)   \\
ob-a         & 61 (7)         & 54 (8)  & 61 (9)   \\
ob-w         & 54 (11)        & 43 (10) & 51 (15)   \\
\end{tabular}
\caption{FYW: best results for object-based approach with majority vote (on the centroids / on the centroids + on the frames) is circa 75\%}
\end{center}
\end{table}

@@

Results: DCASE 2016
